Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.

### 2024-02-05 apples_and_oranges.qmd
several changes

$$SMD_{change} = \frac{Mean_{t_{2}} - Mean_{t_{1}}}{\frac{SD_{t{2}} + SD_{t{1}}}2}\ $$

@@

$$ SE_{change} = \sqrt{\frac{r_{t_{1}t_{2}}}{n} + \frac{SMD^2}{2n}}$$

@@

$$r_{t_{1}t_{2}} \sim \mathcal{TN}(\mu, \sigma, a, b)$$

@@

$$Y_i \sim \mathcal{N}(\theta_i, \sigma_i^2)
$$

$$\theta_i \sim \mathcal{N}(x_i\beta, \tau_i^2)
$$ {#eq-6}

@@

$$Y_i = x_i\beta_i + u_i + \epsilon_i$$ {#eq-7}

@@

$$u_i \sim N(0, \tau^2) $$ {#eq-8}

@@

$$\epsilon_i \sim N(0, \sigma^2)$$ {#eq-9}

@@

$$ \begin{aligned}
    Υ_i &=
    \begin{cases}
        0 & MedControl:  b_0 + u_i + \epsilon_i\\
        1 & MedActive:   b_0 + b_{1_{i}} +  u_i + \epsilon_i \\
        2 & PsyActive:   b_0 + b_{2_{i}}  +  u_i + \epsilon_i \\
        3 & PsyControl:  b_0 + b_{3_{i}}  +  u_i + \epsilon_i \\
    \end{cases}
\end{aligned}$$ {#eq-10}

@@

$$
z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})}
$$ {#eq-11}

@@

Fourth, we tested whether our results could be accounted for by differential regression to the mean. In an initial step, we tested in metaregression as above for differences in baseline means for those medication and psychotherapy studies that shared the same instruments. We then accounted for the effect that possible differing baseline values across all studies studies' outcomes by regressing the SMDs of studies on their respective baseline values of the instrument with which the SMD was calculated. For example, in a study that used the CDRS, this was used to calculate the SMD. This SMD was then used as the outcome and the baseline value of the CDRS, termed *BaselineScore* below, in that study was used as the independent variable as follows: $$
SMD = \beta_0 + \beta_1 BaselineScore + \epsilon
$$ {#eq-12}

This allows us to estimate the effect of the baseline score on the SMD by evaluating $\beta_1$. We can then re-arrange Equation 12 to extract the residuals, that is, the variability in the SMD that is not accounted for by the baseline score.

$$
\epsilon = SMD - (\beta_0 + \beta_1 BaselineScore) ) \
$$

where, $\epsilon$ is the adjusted SMD, for which we simulated standard errors and then used it the outcome in metaregression, as described above for the unadjusted SMDs.

@@

# Plots  -----------------------

#define groups
df_long_for_metan <- df_long_for_metan %>%
  mutate(
    group = case_when(
      psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "Medication Active",
      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "Medication Control",
      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "Psychotherapy Active",
      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "Psychotherapy Control",
      TRUE ~ NA_character_  
    )
  )

group_colors <- c("Medication Active" = "deeppink1", "Medication Control" = "deeppink4", "Psychotherapy Active" = "steelblue1", "Psychotherapy Control" = "steelblue3")

# Plot cohens d per arm

ggplot(df_long_for_metan, aes(x = group, y = cohens_d, color = group)) +
  geom_point(position = position_jitter(width = 0.2), size = 3) +
  geom_segment(x = 4.5, xend = 4.5, y = -3.5, yend = -3, 
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -3, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
  geom_segment(x = 4.5, xend = 4.5, y = -4, yend = -4.5, 
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -4.5, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
  scale_color_manual(values = group_colors) +
  labs(title = "Scatter plot of Cohen's d by group", x = NULL, y = "Cohen's d") +
  theme_minimal() +
  theme(legend.position = "none")

# Now look at change scores of percentage transformed variables

df_long_for_metan <- df_long_for_metan %>% 
  mutate(perc_change = perc_post_mean - perc_baseline_mean)

ggplot(df_long_for_metan, aes(x = group, y = perc_change, color = group)) +
  geom_point(position = position_jitter(width = 0.2), size = 3) +
  geom_segment(x = 4.5, xend = 4.5, y = -0.3, yend = -0.25, 
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -0.3, label = "Less Effective", color = "grey", vjust = 0, hjust = .3) +
  geom_segment(x = 4.5, xend = 4.5, y = -0.4, yend = -0.45, 
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -0.4, label = "More Effective", color = "grey", vjust = 0, hjust = .3) +
  scale_color_manual(values = group_colors) +
  labs(title = "Scatter plot of change scores (percentage transformed) by group", x = NULL, y = "Change") +
  theme_minimal() +
  theme(legend.position = "none")

# # Plot percentage  baseline means 
# 
# ggplot(df_long_for_metan, aes(x = group, y = perc_baseline_mean, color = group)) +
#   geom_point(position = position_jitter(width = 0.2), size = 3) +
#   scale_color_manual(values = group_colors) +
#   labs(title = "Scatter plot of % baseline mean by group", x = NULL, y = "Percentage Baseline Mean") +
#   theme_minimal() +
#   theme(legend.position = "none")

# Box plots

ggplot(df_long_for_metan, aes(x = group, y = cohens_d, fill = group)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(0.8), color = "black", alpha = 0.5) +
  scale_y_continuous(limits = c(-3.5, 1), breaks = seq(-4, 1, by = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
  geom_segment(x = 4.5, xend = 4.5, y = -2.5, yend = -2, 
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -2.2, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
  geom_segment(x = 4.5, xend = 4.5, y = -3, yend = -3.5, 
               arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
  geom_text(x = "Psychotherapy Control", y = -3.2, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
  scale_fill_manual(values = group_colors) +
  labs(title = "Box plot of cohens_d by group", x = "Group", y = "Cohen's d") +
  theme_minimal() +
  theme(legend.position = "none")

@@

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(results_hamd_metareg_baseline[[2]], caption="Table 3 HAM-D baselilne scores")
```

We next addressed potential regression to the mean by residualising SMDs by the baseline scores in one regression model.

@@

# create a new list with dfs for the sims
df_long_for_metan_adj_cohens_d <- df_long_for_metan
# rename the cohens_d value to make space for the new cohen's d
df_long_for_metan_adj_cohens_d <-
  df_long_for_metan_adj_cohens_d %>% 
rename(unadjusted_cohens_d = cohens_d,
) 
# estimate the cohen's d from the residuals
model_adjust_cohens_d <-  lm(df_long_for_metan_adj_cohens_d$unadjusted_cohens_d ~ df_long_for_metan_adj_cohens_d$baseline_mean, na.action=na.exclude)

# there is a great effect of the baseline means, as shown here
results_adjusted_cohens_d <- broom:: tidy(model_adjust_cohens_d )

# extract the residuals which are the adjusted cohen's d, to get the simulated standard errors
df_long_for_metan_adj_cohens_d$cohens_d <- residuals(model_adjust_cohens_d )

# apply function to simulate standard errors
list_df_simulated_adj_cohens_d <- simulate_dataframes_for_st_errors(df = df_long_for_metan_adj_cohens_d, 
                                   num_repetitions = 1000, 
                                   seed  = 1974,
                                   n = length(unique(df_long_for_metan_adj_cohens_d$new_study_id)), 
                                   a = 0.45, 
                                   b = .9, 
                                   mean = 0.65, 
                                   sd = 0.2)

# add the four level variable and re-level
for(i in 1: length(list_df_simulated_adj_cohens_d)){

  list_df_simulated_adj_cohens_d[[i]]$arm_effect_size <- factor(list_df_simulated_adj_cohens_d[[i]]$arm_effect_size)
  list_df_simulated_adj_cohens_d[[i]] <- list_df_simulated_adj_cohens_d[[i]] %>% 
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

  list_df_simulated_adj_cohens_d[[i]]$four_level_var <- factor(list_df_simulated_adj_cohens_d[[i]]$four_level_var) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  #  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
  list_df_simulated_adj_cohens_d[[i]]$four_level_var <-   fct_relevel(list_df_simulated_adj_cohens_d[[i]]$four_level_var,
                                                         "medication_control",
                                                         "medication_active", 
                                                         "psychotherapy_control",
                                                         "psychotherapy_active")
}

# Run metaregression model (assuming that the four level variable is present given that I create this dataframe from the list_)
# specify model
model_1 <- as.formula(~ four_level_var)

# run metareg function 
# here you can use the simulated results directly
list_model_1_meta_reg_adj_cohens_d <- run_metaregression(list_df_simulated_adj_cohens_d, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated_adj_cohens_d[[1]]$four_level_var)
aggregate_results_overall_adjusted_cohens_d <-  aggregate_model_results(list_model_1_meta_reg_adj_cohens_d, condition)
aggregate_results_overall_adjusted_cohens_d <- cbind(aggregate_results_overall_adjusted_cohens_d, condition)

# extract SMDs from the  list
list_dummy_var_means_adjusted_cohens_d <-  lapply(list_model_1_meta_reg_adj_cohens_d, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_from_sim_adjusted_cohens_d <- calculate_mean_coefs_ses(list_dummy_var_means_adjusted_cohens_d, condition)

df_mean_coefs_from_sim_adjusted_cohens_d <- data.frame(cbind(condition, df_mean_coefs_from_sim_adjusted_cohens_d))

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(results_adjusted_cohens_d

              , caption = "Table 4. Effect of baseline values on SMDs")
```

@@

# plot the mean z for the difference between controls
z_psy_con_vs_med_con <- 0
for(i in 1: length(list_model_1_meta_reg_adj_cohens_d )){
  z_psy_con_vs_med_con[i] <- list_model_1_meta_reg_adj_cohens_d [[i]]$zval[condition=="psychotherapy_control"]

stability_sims_adjusted_cohens_d <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)

stab_sims_adjusted_cohens_d <- stability_sims_adjusted_cohens_d %>%
  ggplot(aes(x = n_sims, y = z_value)) +
  geom_point() +
  geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
  geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
               arrow = arrow(length = unit(0.3, "cm")),
               color = "black") +
  annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
  annotate("text", x = 300, y = stability_sims_adjusted_cohens_d$z_value[1] - 0.1, label = "simulated z-values",
           hjust = -0.1, vjust = 0) +
  theme_minimal() +
  labs(
    x = "Number of Simulations",
    y = "z-value",
    title = "Stability of the Statistic of the Difference between \nMedication and Psychotherapy Control Arms",
    subtitle = "Results from 1000 simulations with within-group standard errors"
  )
stab_sims_adjusted_cohens_d
pdf(file = "stab_sims_adjusted_cohens_d.pdf")
stab_sims_adjusted_cohens_d
dev.off()

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
####code to extract baseline-adjusted post means for CDRS and HAMD

de_meaned_baselilne_mean_cdrs <- (list_cdrs_study[[1]]$baseline_mean - mean(list_cdrs_study[[1]]$baseline_mean, na.rm = TRUE))

lm_for_cdrs_smds <- lm(list_cdrs_study[[1]]$post_mean ~ de_meaned_baselilne_mean_cdrs +list_cdrs_study[[1]]$four_level_var,

  na.action=na.exclude,

   )

#summary(lm_for_cdrs_smds)

coefs_lm_for_cdrs_smds  <- broom:: tidy(lm_for_cdrs_smds)

condition <- coefs_lm_for_cdrs_smds$term

cdrs_smds <-c(coefs_lm_for_cdrs_smds$estimate[1], coefs_lm_for_cdrs_smds$estimate[-1] + coefs_lm_for_cdrs_smds$estimate[1])

cdrs_upper_ci <- cdrs_smds + coefs_lm_for_cdrs_smds$std.error*1.96

cdrs_lower_ci <- cdrs_smds - coefs_lm_for_cdrs_smds$std.error*1.96

cdrs_smd_with_cis <- data.frame (

  condition = levels(list_hamd_study$four_level_var),

  cdrs_smds = cdrs_smds[-2],

  cdrs_upper_ci = cdrs_upper_ci[-2],

  cdrs_lower_ci = cdrs_lower_ci[-2]

)

# do the same for hamd

de_meaned_baselilne_mean_hamd <- (list_hamd_study$baseline_mean - mean(list_hamd_study$baseline_mean, na.rm = TRUE))

lm_for_hamd_smds <- lm(list_hamd_study$post_mean ~ de_meaned_baselilne_mean_hamd +list_hamd_study$four_level_var,

                       na.action=na.exclude,

)

#summary(lm_for_hamd_smds)

coefs_lm_for_hamd_smds  <- broom:: tidy(lm_for_hamd_smds)

condition <- coefs_lm_for_hamd_smds$term

hamd_smds <-c(coefs_lm_for_hamd_smds$estimate[1], coefs_lm_for_hamd_smds$estimate[-1] + coefs_lm_for_hamd_smds$estimate[1])

hamd_upper_ci <- hamd_smds + coefs_lm_for_hamd_smds$std.error*1.96

hamd_lower_ci <- hamd_smds - coefs_lm_for_hamd_smds$std.error*1.96

hamd_smd_with_cis <- data.frame (

condition = levels(list_hamd_study$four_level_var),

hamd_smds = hamd_smds[-2],

hamd_upper_ci = hamd_upper_ci[-2],

hamd_lower_ci = hamd_lower_ci[-2]

)

>>>>

$$SMD_{change} = \frac{Mean_{t_{2}} - Mean_{t_{1}}}{\frac{SD_{t{2}} + SD_{t{1}}}2}\ $$ {#eq-4}

@@

$$ SE_{change} = \sqrt{\frac{r_{t_{1}t_{2}}}{n} + \frac{SMD^2}{2n}}$$ {#eq-5}

@@

$$r_{t_{1}t_{2}} \sim \mathcal{TN}(\mu, \sigma, a, b)$$ {#eq-6}

@@

$$Y_i \sim \mathcal{N}(\theta_i, \sigma_i^2)
$$ {#eq-7}

$$\theta_i \sim \mathcal{N}(x_i\beta, \tau_i^2)
$$ {#eq-8}

@@

$$Y_i = x_i\beta_i + u_i + \epsilon_i$$ {#eq-9}

@@

$$u_i \sim N(0, \tau^2) $$ {#eq-10}

@@

$$\epsilon_i \sim N(0, \sigma^2)$$ {#eq-11}

@@

$$ \begin{aligned}
    Υ_i &=
    \begin{cases}
        0 & MedControl:  b_0 + u_i + \epsilon_i\\
        1 & MedActive:   b_0 + b_{1_{i}} +  u_i + \epsilon_i \\
        2 & PsyActive:   b_0 + b_{2_{i}}  +  u_i + \epsilon_i \\
        3 & PsyControl:  b_0 + b_{3_{i}}  +  u_i + \epsilon_i \\
    \end{cases}
\end{aligned}$$ {#eq-12}

@@

$$
z = \frac{\hat{\beta}}{\text{SE}(\hat{\beta})}
$$ {#eq-13}

@@

Fourth, we tested whether our results could be accounted for by differential regression to the mean. In an initial step, we tested in metaregression as above for differences in baseline means. For this we used a min-max normalisation to turn each study arm mean score into a percentage using the following formalism:

$$
\text{outcome}_{percent} = \frac{\bar{X} - \text{scale}_{min}}{\text{scale}_{max} - \text{scale}_{min}}
$$ {#eq-14}

where, $$\bar{X}$$ is the mean score of each study arm on the primary questionnaire, and ${scale}_{min}$ and ${scale}_{max}$ are the minimum and maximum possible value of the scale in question, and for which the the standard deviation is calculated thus:

$$
\text{SD}_{Xpercent} = \frac{\text{SD}_{X}}{\text{scale}_{max} - \text{scale}_{min}}
$$ {#eq-15}

where $\text{SD}_{X}$ is the original standard deviation of the mean for X. Applying equations (12) and (13) on the baseline and outcome means and standard deviations of each study allowed us to then calculate SMDs and standard errors using equations (4) and (5)

@@

# Plots  -----------------------
# 
# #define groups
# df_long_for_metan <- df_long_for_metan %>%
#   mutate(
#     group = case_when(
#       psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "Medication Active",
#       psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "Medication Control",
#       psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "Psychotherapy Active",
#       psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "Psychotherapy Control",
#       TRUE ~ NA_character_  
#     )
#   )

# group_colors <- c("Medication Active" = "deeppink1", "Medication Control" = "deeppink4", "Psychotherapy Active" = "steelblue1", "Psychotherapy Control" = "steelblue3")
# 
# # Plot cohens d per arm
# 
# ggplot(df_long_for_metan, aes(x = group, y = cohens_d, color = group)) +
#   geom_point(position = position_jitter(width = 0.2), size = 3) +
#   geom_segment(x = 4.5, xend = 4.5, y = -3.5, yend = -3, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -3, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
#   geom_segment(x = 4.5, xend = 4.5, y = -4, yend = -4.5, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -4.5, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
#   scale_color_manual(values = group_colors) +
#   labs(title = "Scatter plot of Cohen's d by group", x = NULL, y = "Cohen's d") +
#   theme_minimal() +
#   theme(legend.position = "none")
# 
# # Now look at change scores of percentage transformed variables
# 
# df_long_for_metan <- df_long_for_metan %>% 
#   mutate(perc_change = perc_post_mean - perc_baseline_mean)
# 
# ggplot(df_long_for_metan, aes(x = group, y = perc_change, color = group)) +
#   geom_point(position = position_jitter(width = 0.2), size = 3) +
#   geom_segment(x = 4.5, xend = 4.5, y = -0.3, yend = -0.25, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -0.3, label = "Less Effective", color = "grey", vjust = 0, hjust = .3) +
#   geom_segment(x = 4.5, xend = 4.5, y = -0.4, yend = -0.45, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -0.4, label = "More Effective", color = "grey", vjust = 0, hjust = .3) +
#   scale_color_manual(values = group_colors) +
#   labs(title = "Scatter plot of change scores (percentage transformed) by group", x = NULL, y = "Change") +
#   theme_minimal() +
#   theme(legend.position = "none")
# 
# # # Plot percentage  baseline means 
# # 
# # ggplot(df_long_for_metan, aes(x = group, y = perc_baseline_mean, color = group)) +
# #   geom_point(position = position_jitter(width = 0.2), size = 3) +
# #   scale_color_manual(values = group_colors) +
# #   labs(title = "Scatter plot of % baseline mean by group", x = NULL, y = "Percentage Baseline Mean") +
# #   theme_minimal() +
# #   theme(legend.position = "none")
# 
# # Box plots
# 
# ggplot(df_long_for_metan, aes(x = group, y = cohens_d, fill = group)) +
#   geom_boxplot(outlier.shape = NA, position = position_dodge(0.8), color = "black", alpha = 0.5) +
#   scale_y_continuous(limits = c(-3.5, 1), breaks = seq(-4, 1, by = 1)) +
#   geom_hline(yintercept = 0, linetype = "dashed", color = "grey", size = 1) +
#   geom_segment(x = 4.5, xend = 4.5, y = -2.5, yend = -2, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -2.2, label = "Less Effective", color = "grey", vjust = 0, hjust = .2) +
#   geom_segment(x = 4.5, xend = 4.5, y = -3, yend = -3.5, 
#                arrow = arrow(length = unit(0.25, "cm"), type = "closed"), color = "grey") +
#   geom_text(x = "Psychotherapy Control", y = -3.2, label = "More Effective", color = "grey", vjust = 0, hjust = .2) +
#   scale_fill_manual(values = group_colors) +
#   labs(title = "Box plot of cohens_d by group", x = "Group", y = "Cohen's d") +
#   theme_minimal() +
#   theme(legend.position = "none")

@@

### do ALL studies using percentage at baseline
all_studies_baseline_metareg <- list_df_simulated_percentage[[1]]

#turn sds to sec
all_studies_baseline_metareg$baseline_stand_error <- all_studies_baseline_metareg$perc_baseline_sd/
                                        sqrt(all_studies_baseline_metareg$baseline_n)

model_1 <- as.formula(~ four_level_var)
condition <-  levels(list_df_simulated[[1]]$four_level_var) # this is the four level var 
                                                              #throughout
# also change this for the function to work
all_studies_baseline_metareg$baseline_mean <- all_studies_baseline_metareg$perc_baseline_mean
results_all_studies_metareg_baseline <- run_metareg_and_extract(all_studies_baseline_metareg,  
                                                         model_1, condition)

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
knitr:: kable(results_hamd_metareg_baseline[[2]], caption="Table 3b HAM-D baselilne scores")
```

We next addressed the e

@@

# # create a new list with dfs for the sims
# df_long_for_metan_adj_cohens_d <- df_long_for_metan
# # rename the cohens_d value to make space for the new cohen's d
# df_long_for_metan_adj_cohens_d <-
#   df_long_for_metan_adj_cohens_d %>% 
# rename(unadjusted_cohens_d = cohens_d,
# ) 
# # estimate the cohen's d from the residuals
# model_adjust_cohens_d <-  lm(df_long_for_metan_adj_cohens_d$unadjusted_cohens_d ~ df_long_for_metan_adj_cohens_d$baseline_mean, na.action=na.exclude)
# 
# # there is a great effect of the baseline means, as shown here
# results_adjusted_cohens_d <- broom:: tidy(model_adjust_cohens_d )
# 
# # extract the residuals which are the adjusted cohen's d, to get the simulated standard errors
# df_long_for_metan_adj_cohens_d$cohens_d <- residuals(model_adjust_cohens_d )
# 
# # apply function to simulate standard errors
# list_df_simulated_adj_cohens_d <- simulate_dataframes_for_st_errors(df = df_long_for_metan_adj_cohens_d, 
#                                    num_repetitions = 1000, 
#                                    seed  = 1974,
#                                    n = length(unique(df_long_for_metan_adj_cohens_d$new_study_id)), 
#                                    a = 0.45, 
#                                    b = .9, 
#                                    mean = 0.65, 
#                                    sd = 0.2)
# 
# 
# # add the four level variable and re-level
# for(i in 1: length(list_df_simulated_adj_cohens_d)){
#   
#   list_df_simulated_adj_cohens_d[[i]]$arm_effect_size <- factor(list_df_simulated_adj_cohens_d[[i]]$arm_effect_size)
#   list_df_simulated_adj_cohens_d[[i]] <- list_df_simulated_adj_cohens_d[[i]] %>% 
#     mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
#                                       psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
#                                       psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
#                                       psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
#     )
#   
#   list_df_simulated_adj_cohens_d[[i]]$four_level_var <- factor(list_df_simulated_adj_cohens_d[[i]]$four_level_var) # turn to factor
#   
#   # relevel so that medication control becomes the reference category for the regression
#   #  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
#   list_df_simulated_adj_cohens_d[[i]]$four_level_var <-   fct_relevel(list_df_simulated_adj_cohens_d[[i]]$four_level_var,
#                                                          "medication_control",
#                                                          "medication_active", 
#                                                          "psychotherapy_control",
#                                                          "psychotherapy_active")
# }
# 
# 
# 
# 
# 
# # Run metaregression model (assuming that the four level variable is present given that I create this dataframe from the list_)
# # specify model
# model_1 <- as.formula(~ four_level_var)
# 
# # run metareg function 
# # here you can use the simulated results directly
# list_model_1_meta_reg_adj_cohens_d <- run_metaregression(list_df_simulated_adj_cohens_d, model_1)
# 
# # extract coefficients and model characteristics.
# condition <- levels(list_df_simulated_adj_cohens_d[[1]]$four_level_var)
# aggregate_results_overall_adjusted_cohens_d <-  aggregate_model_results(list_model_1_meta_reg_adj_cohens_d, condition)
# aggregate_results_overall_adjusted_cohens_d <- cbind(aggregate_results_overall_adjusted_cohens_d, condition)
# 
# # extract SMDs from the  list
# list_dummy_var_means_adjusted_cohens_d <-  lapply(list_model_1_meta_reg_adj_cohens_d, extract_coefficients_func)
# 
# # calculate mean SMDs and ses
# df_mean_coefs_from_sim_adjusted_cohens_d <- calculate_mean_coefs_ses(list_dummy_var_means_adjusted_cohens_d, condition)
# 
# df_mean_coefs_from_sim_adjusted_cohens_d <- data.frame(cbind(condition, df_mean_coefs_from_sim_adjusted_cohens_d))
# 
# 
# 
#

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# knitr:: kable(results_adjusted_cohens_d 
#               
#               , caption = "Table 4. Effect of baseline values on SMDs")
```

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# 
# # plot the mean z for the difference between controls
# z_psy_con_vs_med_con <- 0
# for(i in 1: length(list_model_1_meta_reg_adj_cohens_d )){
#   z_psy_con_vs_med_con[i] <- list_model_1_meta_reg_adj_cohens_d [[i]]$zval[condition=="psychotherapy_control"]
#   
# }
# 
# 
# 
# stability_sims_adjusted_cohens_d <- data.frame(n_sims = 1:1000, z_value = z_psy_con_vs_med_con)
# 
# 
# stab_sims_adjusted_cohens_d <- stability_sims_adjusted_cohens_d %>%
#   ggplot(aes(x = n_sims, y = z_value)) +
#   geom_point() +
#   geom_hline(yintercept = 1.65, linetype = "dotted", color = "red") +
#   geom_segment(aes(x = 300, xend = 300, y = 2, yend = 1.65),
#                arrow = arrow(length = unit(0.3, "cm")),
#                color = "black") +
#   annotate("text", x = 300, y = 2.0, label = "p < 0.05 threshold (values above line significant)", hjust = -0.1, vjust = 0) +
#   annotate("text", x = 300, y = stability_sims_adjusted_cohens_d$z_value[1] - 0.1, label = "simulated z-values",
#            hjust = -0.1, vjust = 0) +
#   theme_minimal() +
#   labs(
#     x = "Number of Simulations",
#     y = "z-value",
#     title = "Stability of the Statistic of the Difference between \nMedication and Psychotherapy Control Arms",
#     subtitle = "Results from 1000 simulations with within-group standard errors"
#   )
# stab_sims_adjusted_cohens_d
# pdf(file = "stab_sims_adjusted_cohens_d.pdf")
# stab_sims_adjusted_cohens_d
# dev.off()
#

@@

```{# {r, warning=FALSE, message = FALSE, echo=FALSE}
# ####code to extract baseline-adjusted post means for CDRS and HAMD 
# 
#  
# 
# de_meaned_baselilne_mean_cdrs <- (list_cdrs_study[[1]]$baseline_mean - mean(list_cdrs_study[[1]]$baseline_mean, na.rm = TRUE)) 
# 
#                                     
# 
#  
# 
#                                  
# 
# lm_for_cdrs_smds <- lm(list_cdrs_study[[1]]$post_mean ~ de_meaned_baselilne_mean_cdrs +list_cdrs_study[[1]]$four_level_var,
# 
#  
# 
#   na.action=na.exclude,
# 
#    )
# 
#  
# 
#  
# 
# #summary(lm_for_cdrs_smds)
# 
#  
# 
# coefs_lm_for_cdrs_smds  <- broom:: tidy(lm_for_cdrs_smds)
# 
#  
# 
# condition <- coefs_lm_for_cdrs_smds$term
# 
# cdrs_smds <-c(coefs_lm_for_cdrs_smds$estimate[1], coefs_lm_for_cdrs_smds$estimate[-1] + coefs_lm_for_cdrs_smds$estimate[1])
# 
# cdrs_upper_ci <- cdrs_smds + coefs_lm_for_cdrs_smds$std.error*1.96
# 
# cdrs_lower_ci <- cdrs_smds - coefs_lm_for_cdrs_smds$std.error*1.96
# 
#  
# 
# cdrs_smd_with_cis <- data.frame (
# 
#   condition = levels(list_hamd_study$four_level_var),
# 
#   cdrs_smds = cdrs_smds[-2],
# 
#   cdrs_upper_ci = cdrs_upper_ci[-2],
# 
#   cdrs_lower_ci = cdrs_lower_ci[-2]
# 
# )
# 
#  
# 
#  
# 
# # do the same for hamd
# 
# de_meaned_baselilne_mean_hamd <- (list_hamd_study$baseline_mean - mean(list_hamd_study$baseline_mean, na.rm = TRUE)) 
# 
#  
# 
#  
# 
#  
# 
# lm_for_hamd_smds <- lm(list_hamd_study$post_mean ~ de_meaned_baselilne_mean_hamd +list_hamd_study$four_level_var,
# 
#                       
# 
#                        na.action=na.exclude,
# 
# )
# 
#  
# 
#  
# 
# #summary(lm_for_hamd_smds)
# 
#  
# 
# coefs_lm_for_hamd_smds  <- broom:: tidy(lm_for_hamd_smds)
# 
#  
# 
#  
# 
# condition <- coefs_lm_for_hamd_smds$term
# 
# hamd_smds <-c(coefs_lm_for_hamd_smds$estimate[1], coefs_lm_for_hamd_smds$estimate[-1] + coefs_lm_for_hamd_smds$estimate[1])
# 
# hamd_upper_ci <- hamd_smds + coefs_lm_for_hamd_smds$std.error*1.96
# 
# hamd_lower_ci <- hamd_smds - coefs_lm_for_hamd_smds$std.error*1.96
# 
#  
# 
# hamd_smd_with_cis <- data.frame (
# 
# condition = levels(list_hamd_study$four_level_var),
# 
# hamd_smds = hamd_smds[-2],
# 
# hamd_upper_ci = hamd_upper_ci[-2],
# 
# hamd_lower_ci = hamd_lower_ci[-2]
# 
# )
# 
#  
# 
#
